\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{times}
\usepackage{setspace}
\usepackage{enumitem}
\onehalfspacing
\setlength{\parindent}{0pt}
\setlength{\parskip}{4pt}

\begin{document}

\begin{center}
\textbf{\large Individual Contribution Report --- Project 2}\\[4pt]
TransferDB Application --- CMPE 321, Spring 2026
\end{center}

\vspace{2pt}

\textbf{Personal Information.}
Name: Ahmet Akda\u{g} \quad Student ID: 2020400057 \quad Group Number: 01

\textbf{Tasks \& Contributions.}
I owned the backend implementation of the TransferDB application. On
the database side I wrote the complete \texttt{init.sql}: I extended
the Project 1 schema with a \texttt{Suspension} table and a
\texttt{Club.city} column, authored all seven triggers that enforce
the business rules that SQL \texttt{CHECK} cannot express
(120-minute scheduling gap on insert, attendance $\leq$ stadium
capacity on insert and update, $\leq$11 starters / $\leq$23 squad per
club per match, $\geq$11 players at result-submission time, the
loan-vs-parent-club ban, the suspension block against carded players,
C8 city consistency, and C9 no overlapping contracts of the same
type), and authored the two stored procedures. The transfer procedure
auto-terminates the player's previous Permanent contract, writes the
new Contract and Transfer\_Record, and updates \texttt{market\_value}
for Purchase transfers. The match-result procedure applies the forum
\#4 reset semantics: it inserts Suspension rows for red cards and for
5-yellow accumulations (with the window bounded to matches strictly
after the player's last served suspension in the same competition),
and marks prior suspensions as served when the player was kept out of
the squad of their club's next match. I also added the
\texttt{@DISABLE\_TRIGGERS} session-variable bypass so the
intentionally-invalid seed records described in spec \S2.3 can be
loaded without tripping the validation triggers, while any direct
MySQL client write outside the seed path remains fully protected.

On the application side I wrote \texttt{app.py} as a single Flask
module covering all 18 key operations for the four roles, with
\texttt{@current\_user\_id}/\texttt{@current\_user\_role} session
variables pushed on every database connection so the referee-only
match-result rule survives even when the web interface is bypassed
from a MySQL client. I authored every SQL query manually (no ORM),
used \texttt{\%s} parameterized placeholders everywhere for injection
safety, and implemented the password policy and bcrypt hashing. I
wrote all 22 Jinja templates (role-aware navigation, per-role
dashboards, match squad selection, match-result entry, fixtures,
standings, squad statistics, competition leaderboards, career
history, and the DB Manager CRUD forms). I rewrote
\texttt{import\_users.py} to ingest all eleven
\texttt{initial\_data.xlsx} sheets in FK-safe order with NaN
coercion, and I brought up the MySQL container under Docker to
smoke-test every page end-to-end and to verify each trigger
rejection via direct \texttt{mysql} client.

\textbf{Collaboration \& Challenges.}
We split the work cleanly: I owned implementation and debugging, my
teammate owned the written analysis in Part 3 and the testing
scenarios against the spec. We synced through the shared repository
and reconciled whenever a schema decision affected the normalization
write-up. The hardest part was the suspension lifecycle: MySQL
forbids a trigger from modifying the same table it fires on, which
pushed the auto-termination of the previous Permanent contract into
a stored procedure rather than a trigger, and the forum \#4 reset
rule required a correctly-windowed yellow-card sum subquery.

\textbf{Self-Assessment.}
I strengthened my understanding of trigger semantics, of how to
propagate authentication context into the database through session
variables, and of where the boundary between
\texttt{CHECK}-expressible and trigger-expressible constraints
actually lies. Next time I would invest earlier in integration tests
that call the stored procedures directly from a MySQL client rather
than only through the web app.

\textbf{Use of AI Tools.}
No AI tools were used for this project. All code, SQL, and
documentation were written by the team.

\end{document}
