\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{array}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage{amssymb}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}

\title{TransferDB --- Part 3:\\Schema Refinement and Normalization}
\author{Group 01}
\date{April 2026}

\begin{document}
\maketitle

\section{Introduction}

This document analyzes the relational schema from Project 1 for
normalization compliance, catalogs every non-trivial functional
dependency (FD) derived strictly from the project rules (forum
clarification \#2), and reports the additional integrity constraints
that were moved into \texttt{CHECK} constraints, triggers, and stored
procedures in \texttt{init.sql}.

Throughout, \emph{non-trivial} means an FD $X \to Y$ where $Y \not\subseteq X$.

\section{Functional Dependencies}

FDs are enumerated below per relation. Each FD follows from an explicit
rule in the project specification or an explicit uniqueness/primary-key
constraint. No FD is added from ``real-world'' assumptions, per forum
clarification \#2.

\subsection*{Person}
\begin{itemize}[leftmargin=*]
  \item \texttt{person\_ID} $\to$ \texttt{name, surname, nationality, date\_of\_birth} \hfill (PK)
\end{itemize}

\subsection*{Player}
\begin{itemize}[leftmargin=*]
  \item \texttt{person\_ID} $\to$ \texttt{username, password, market\_value, main\_position, strong\_foot, height} \hfill (PK)
  \item \texttt{username} $\to$ \texttt{person\_ID} \hfill (UNIQUE)
\end{itemize}

\subsection*{Manager}
\begin{itemize}[leftmargin=*]
  \item \texttt{person\_ID} $\to$ \texttt{username, password, preferred\_formation, experience\_level} \hfill (PK)
  \item \texttt{username} $\to$ \texttt{person\_ID} \hfill (UNIQUE)
\end{itemize}

\subsection*{Referee}
\begin{itemize}[leftmargin=*]
  \item \texttt{person\_ID} $\to$ \texttt{username, password, license\_level, years\_of\_experience} \hfill (PK)
  \item \texttt{username} $\to$ \texttt{person\_ID} \hfill (UNIQUE)
\end{itemize}

\subsection*{DB\_Manager}
\begin{itemize}[leftmargin=*]
  \item \texttt{username} $\to$ \texttt{password} \hfill (PK)
\end{itemize}

\subsection*{Stadium}
\begin{itemize}[leftmargin=*]
  \item \texttt{stadium\_ID} $\to$ \texttt{stadium\_name, city, capacity} \hfill (PK)
  \item \texttt{(stadium\_name, city)} $\to$ \texttt{stadium\_ID, capacity} \hfill (UNIQUE)
\end{itemize}

\subsection*{Club}
\begin{itemize}[leftmargin=*]
  \item \texttt{club\_ID} $\to$ \texttt{club\_name, stadium\_ID, city, foundation\_year, manager\_ID} \hfill (PK)
  \item \texttt{club\_name} $\to$ \texttt{club\_ID, \dots} \hfill (UNIQUE)
  \item \texttt{manager\_ID} $\to$ \texttt{club\_ID, \dots} \hfill (\emph{Manages} is 1:1)
  \item \texttt{stadium\_ID} $\to$ \texttt{city} \hfill (derived from rule C8:
    club city equals its home stadium's city, so knowing the stadium fixes
    the city regardless of which club we look at)
\end{itemize}

\subsection*{Contract}
\begin{itemize}[leftmargin=*]
  \item \texttt{contract\_id} $\to$ \texttt{player\_id, club\_id, start\_date, end\_date, weekly\_wage, contract\_type} \hfill (PK)
\end{itemize}

\subsection*{Transfer\_Record}
\begin{itemize}[leftmargin=*]
  \item \texttt{transfer\_id} $\to$ \texttt{player\_id, from\_club\_id, to\_club\_id, transfer\_date, transfer\_fee, transfer\_type} \hfill (PK)
\end{itemize}

\subsection*{Competition}
\begin{itemize}[leftmargin=*]
  \item \texttt{competition\_ID} $\to$ \texttt{name, season, country, competition\_type} \hfill (PK)
  \item \texttt{(name, season)} $\to$ \texttt{competition\_ID, country, competition\_type} \hfill (UNIQUE)
\end{itemize}

\subsection*{Match}
\begin{itemize}[leftmargin=*]
  \item \texttt{match\_ID} $\to$ \texttt{competition\_ID, home\_club\_ID, away\_club\_ID, stadium\_ID, referee\_ID, match\_datetime, attendance, home\_goals, away\_goals} \hfill (PK)
  \item \texttt{(stadium\_ID, match\_datetime)} $\to$ all other attributes \hfill (forum \#2: 120-min rule forces uniqueness)
  \item \texttt{(referee\_ID, match\_datetime)} $\to$ all other attributes \hfill (120-min rule for referees)
  \item \texttt{(home\_club\_ID, match\_datetime)} $\to$ all other attributes \hfill (120-min rule for clubs)
  \item \texttt{(away\_club\_ID, match\_datetime)} $\to$ all other attributes \hfill (120-min rule for clubs)
\end{itemize}

\subsection*{Match\_Stats}
\begin{itemize}[leftmargin=*]
  \item \texttt{(match\_ID, player\_id)} $\to$ \texttt{club\_id, is\_starter, minutes\_played,
    position\_played, goals, assists, yellow\_cards, red\_cards, rating} \hfill (PK)
  \item \texttt{(match\_ID, player\_id)} $\to$ \texttt{club\_id} is reinforced
    by forum clarification C7: a player appears for exactly one side per match.
\end{itemize}

\subsection*{Suspension}
\begin{itemize}[leftmargin=*]
  \item \texttt{suspension\_id} $\to$ all other attributes \hfill (PK)
  \item \texttt{(player\_id, triggering\_match\_ID)} $\to$ \texttt{suspension\_id, competition\_ID, reason, status, served\_match\_ID} \hfill (UNIQUE)
\end{itemize}

\section{BCNF Analysis}

A relation is in Boyce-Codd Normal Form (BCNF) if, for every non-trivial
FD $X \to Y$, $X$ is a superkey. We verify each relation below.

\begin{longtable}{@{}lp{0.75\linewidth}@{}}
\toprule
\textbf{Relation} & \textbf{Analysis} \\
\midrule\endhead

Person & Only FD is \texttt{person\_ID} $\to$ rest. PK is a superkey. \textbf{BCNF.} \\[2pt]

Player  & All LHSs (\texttt{person\_ID}, \texttt{username}) are
  keys (PK and alternate key). \textbf{BCNF.} \\[2pt]

Manager & Same structure as Player. \textbf{BCNF.} \\[2pt]

Referee & Same structure as Player. \textbf{BCNF.} \\[2pt]

DB\_Manager & Only FD \texttt{username} $\to$ \texttt{password}; LHS is PK.
  \textbf{BCNF.} \\[2pt]

Stadium & LHSs: \texttt{stadium\_ID} (PK), \texttt{(stadium\_name,city)}
  (UNIQUE). Both superkeys. \textbf{BCNF.} \\[2pt]

Club & LHSs: \texttt{club\_ID} (PK), \texttt{club\_name} (UNIQUE),
  \texttt{manager\_ID} (UNIQUE, Manages is 1:1). \emph{However,}
  \texttt{stadium\_ID} $\to$ \texttt{city} (rule C8) is a non-trivial FD
  whose LHS is \emph{not} a superkey of Club --- multiple clubs may share
  one stadium, so \texttt{stadium\_ID} is not unique in Club. This causes
  a BCNF violation. See decomposition discussion below. \\[2pt]

Contract & Only FD: \texttt{contract\_id} $\to$ rest. LHS is PK.
  \textbf{BCNF.} \\[2pt]

Transfer\_Record & PK $\to$ rest only. \textbf{BCNF.} \\[2pt]

Competition & LHSs: \texttt{competition\_ID} (PK),
  \texttt{(name, season)} (UNIQUE). Both superkeys. \textbf{BCNF.} \\[2pt]

Match & LHSs include \texttt{match\_ID} (PK) and the four
  \texttt{(X, match\_datetime)} candidate keys from forum \#2. All are
  superkeys. \textbf{BCNF.} \\[2pt]

Match\_Stats & Only non-trivial FD is
  \texttt{(match\_ID, player\_id)} $\to$ rest. LHS is PK. \textbf{BCNF.}
  \\[2pt]

Suspension & LHSs: \texttt{suspension\_id} (PK) and
  \texttt{(player\_id, triggering\_match\_ID)} (UNIQUE). Both superkeys.
  \textbf{BCNF.} \\

\bottomrule
\end{longtable}

\subsection{Club: BCNF Discussion and Our Decision}

The FD \texttt{stadium\_ID} $\to$ \texttt{city} in Club is the only
BCNF-violating dependency in the schema. Because multiple clubs may
share a stadium (the spec's \emph{Home\_Stadium} relationship is
\texttt{Club $\to$ Stadium}, N:1), \texttt{stadium\_ID} is not a key of
Club, so the dependency places the same fact
(\texttt{stadium\_ID} $\to$ \texttt{city}) in many Club rows
redundantly.

\textbf{3NF check:} The dependency is trivial in 3NF terms because
\texttt{city} is already functionally dependent on a superkey
(\texttt{club\_ID}). Since \texttt{city} is also determined (indirectly)
by \texttt{stadium\_ID}, the \texttt{city} column is \emph{partially
redundant} but Club is still in 3NF (the condition ``every non-prime
attribute is non-transitively dependent on every candidate key'' holds
in the weaker 3NF formulation because the chain
\texttt{club\_ID} $\to$ \texttt{stadium\_ID} $\to$ \texttt{city} is
exactly the kind of transitive dependency that a BCNF decomposition
would remove).

\textbf{Candidate decompositions:}
\begin{enumerate}[leftmargin=*]
\item \emph{Drop \texttt{city} from Club.} Derive it via a join to
  Stadium at read time. This gives
  \texttt{Club(club\_ID, club\_name, stadium\_ID, foundation\_year, manager\_ID)}
  which is in BCNF and lossless-join / dependency-preserving (Stadium
  already holds city).
\item \emph{Project into two relations:}
  \texttt{Club1(club\_ID, club\_name, stadium\_ID, foundation\_year, manager\_ID)}
  and \texttt{Club2(stadium\_ID, city)} (which is a restriction of
  Stadium). Still BCNF, lossless, dependency-preserving, but redundant
  with Stadium.
\end{enumerate}

\textbf{Our decision:} we kept \texttt{city} inside Club rather than
decomposing. The driving reason is that rule C8 (``club city equals
stadium city'') must be \emph{enforced} at write time. With the
\texttt{city} column physically present in Club we can validate it in
triggers \texttt{trg\_club\_city\_ins} and
\texttt{trg\_club\_city\_upd} without cross-row joins in the trigger
body. Removing the column would shift validation to the application
layer or force each write to re-check via a subquery against Stadium,
which MySQL \texttt{CHECK} cannot do and which is more expensive per
insert. The schema is in 3NF, and the redundancy is compensated for by
the two triggers, which guarantee that the \texttt{city} value in Club
is always identical to its home stadium's.

This is a deliberate trade-off between BCNF purity and enforceable
integrity: since the ``anomaly'' that BCNF wants to prevent (someone
updating Stadium.city without updating Club.city) is exactly the
anomaly the trigger prevents, the practical gain of decomposition is
zero.

All other relations are in BCNF without decomposition.

\section{Added Constraints Beyond Project 1 DDL}

Project 1's DDL captured what SQL \texttt{CHECK} can express locally.
Project 2 added every remaining business rule at the database level
through triggers and stored procedures, so the database rejects invalid
writes regardless of whether they come from the web UI or a direct
MySQL client (spec \S 4, ``Strict Enforcement Rule'').

\subsection{Local \texttt{CHECK} constraints (from Project 1 and tightened)}

\begin{itemize}[leftmargin=*]
  \item \texttt{Player.market\_value} $\geq 0$, \texttt{height} $> 0$.
  \item \texttt{Referee.years\_of\_experience} $\geq 0$.
  \item \texttt{Stadium.capacity} $> 0$ and \texttt{(stadium\_name, city)} unique.
  \item \texttt{Contract.weekly\_wage} $> 0$, \texttt{end\_date} $>$ \texttt{start\_date}.
  \item \texttt{Transfer\_Record.transfer\_fee} $\geq 0$;
    \texttt{from\_club\_id} $\neq$ \texttt{to\_club\_id};
    \texttt{Free} $\Rightarrow$ \texttt{fee} $= 0$
    and non-\texttt{Free} $\Rightarrow$ \texttt{fee} $> 0$.
  \item \texttt{Match.home\_club\_ID} $\neq$ \texttt{away\_club\_ID};
    non-negative scores / attendance.
  \item \texttt{Match\_Stats}: \texttt{minutes\_played} $\in [0, 120]$,
    yellow \& red card ranges, \texttt{yellow\_cards < 2 OR red\_cards = 1}
    (implements ``2 yellows $\Rightarrow$ red''),
    \texttt{rating} $\in [1.0, 10.0]$.
  \item \texttt{Competition.competition\_type} is a strict
    \texttt{ENUM('League','Cup','International')} (forum \#1).
\end{itemize}

\subsection{Trigger-enforced constraints (Project 2 additions)}

These require cross-row or cross-table reasoning that \texttt{CHECK}
cannot express.

\begin{longtable}{@{}p{0.25\linewidth}p{0.35\linewidth}p{0.35\linewidth}@{}}
\toprule
\textbf{Rule} & \textbf{Trigger / procedure} & \textbf{Enforcement} \\
\midrule\endhead

C1 --- Loan needs covering Permanent at a different club &
  \texttt{trg\_contract\_ins} &
  Counts rows in Contract where \texttt{contract\_type='Permanent'},
  \texttt{club\_id}$\neq$ new club, \texttt{start}$\leq$ new start,
  \texttt{end}$\geq$ new end. Reject if 0. \\

C2 --- 120-minute gap between matches sharing stadium/club/ref &
  \texttt{trg\_match\_ins} &
  Scans existing Matches with overlapping predicate and
  \texttt{ABS(TIMESTAMPDIFF(MINUTE, \dots)) < 120}. Reject if any. \\

C3 --- Attendance $\leq$ stadium capacity &
  \texttt{trg\_match\_ins}, \texttt{trg\_match\_upd} &
  Joins to Stadium, compares capacity. Fires on insert and on any
  later update of attendance. \\

C4a --- $\leq$ 11 starters per club per match &
  \texttt{trg\_match\_stats\_ins} &
  Counts current starters for this (match, club) before allowing a
  new starter row. \\

C4b --- $\leq$ 23 total players per club per match &
  \texttt{trg\_match\_stats\_ins} &
  Similar count on the whole squad. \\

C4c --- $\geq$ 11 players per club at result time &
  \texttt{trg\_match\_upd} &
  When \texttt{home\_goals} transitions \texttt{NULL} $\to$ set, counts
  both sides; rejects if either has fewer than 11 squad members. \\

C5a --- Red card $\Rightarrow$ next-match suspension &
  \texttt{sp\_submit\_match\_result} &
  After updating Match, inserts into \texttt{Suspension} for every
  player with \texttt{red\_cards=1}. \\

C5b --- 5 yellow cards in a competition $\Rightarrow$ 1-match ban
  (count resets after ban, per forum \#4) &
  \texttt{sp\_submit\_match\_result} &
  Sums \texttt{yellow\_cards} for each player since the max
  \texttt{served\_match\_ID} datetime for that player in that
  competition; if $\geq 5$, creates a \texttt{5-yellow} Suspension
  (or upgrades an existing \texttt{red} entry to \texttt{combined}). \\

C5c --- Suspended players cannot appear in the next match's squad &
  \texttt{trg\_match\_stats\_ins} &
  Rejects insert if a \texttt{pending} Suspension exists for this
  (player, competition) triggered by an earlier match. \\

C5d --- Serving a ban consumes it and resets the yellow window
  (forum \#4) &
  \texttt{sp\_submit\_match\_result} &
  After the match is played, marks pending suspensions for
  players-of-this-match's-clubs as \texttt{served} when the player was
  not in the squad. The 5-yellow subquery then re-counts only yellows
  whose match datetime exceeds the last served suspension. \\

C6a --- Free transfer fee $= 0$ &
  \texttt{CHECK} &
  Direct \texttt{CHECK} constraint. \\

C6b --- Loan transfer \texttt{from\_club} equals parent's Permanent
  club &
  \texttt{trg\_transfer\_ins} &
  Resolves the player's active Permanent club at transfer date; reject
  if \texttt{from\_club\_id} differs. \\

C7a --- Player's \texttt{club\_id} must match home or away of the
  match &
  \texttt{trg\_match\_stats\_ins} &
  Compares against Match row. \\

C7b --- Loan player cannot appear for parent club &
  \texttt{trg\_match\_stats\_ins} &
  If player has an active Loan, resolves parent Permanent club and
  rejects if the new row's \texttt{club\_id} equals the parent. \\

C8 --- Club city = Stadium city &
  \texttt{trg\_club\_city\_ins}, \texttt{trg\_club\_city\_upd} &
  Reads Stadium.city and compares to the incoming Club.city. \\

C9 --- No overlapping active contracts of the same type &
  \texttt{trg\_contract\_ins} &
  Strict date-range overlap check per player + type. \\

Future-dated match scheduling &
  \texttt{trg\_match\_ins} &
  \texttt{NEW.match\_datetime > NOW()}. \\

Only assigned referee may submit results &
  \texttt{trg\_match\_upd} &
  Checks MySQL session variables \texttt{@current\_user\_id} and
  \texttt{@current\_user\_role}, which the Flask app sets on every
  connection after login. Direct MySQL-client writes (no session vars)
  are rejected. \\

Match results only after scheduled time &
  \texttt{trg\_match\_upd} &
  Refuses the \texttt{NULL} $\to$ filled transition when
  \texttt{match\_datetime > NOW()}. \\

At most one active Permanent per player + auto-termination &
  \texttt{trg\_contract\_ins} + \texttt{sp\_register\_transfer} &
  Trigger forbids a second active Permanent; the procedure terminates
  the old Permanent (setting \texttt{end\_date} to today) before
  writing the new one. \\

Purchase transfer updates \texttt{Player.market\_value} &
  \texttt{sp\_register\_transfer} &
  After writing the Transfer\_Record and Contract, sets
  \texttt{market\_value = transfer\_fee}. \\

\bottomrule
\end{longtable}

\subsection{Password and SQL-injection safeguards (application layer)}

Per spec \S 4, password hashing (bcrypt), password policy validation
(length/complexity), and parameterized queries are handled in the
Python application layer, not in the database. All DB writes go through
\texttt{mysql-connector}'s \texttt{\%s} placeholders, never through
string concatenation.

\section{Forum Clarifications Applied}

\begin{itemize}[leftmargin=*]
  \item \textbf{\#1 (Competition type strict enum):}
    \texttt{competition\_type} is an \texttt{ENUM} of three values; the
    DB Manager ``create competition'' form only exposes these three
    options.
  \item \textbf{\#2 (FD listing):} Only FDs derived from explicit rules
    are listed, including the \texttt{(stadium\_id, match\_datetime)}
    family that is forced into existence by the 120-min rule.
  \item \textbf{\#3 (No suspension carry-over across seasons):}
    Suspensions are keyed by \texttt{competition\_ID}, which uniquely
    identifies a (name, season) pair; queries never unify across
    seasons.
  \item \textbf{\#4 (Red $+$ 5-yellow concurrent = 1-match ban; yellow
    count resets after a served ban):} the 5-yellow subquery in
    \texttt{sp\_submit\_match\_result} windows yellows strictly after
    the player's last \texttt{served} Suspension in the same
    competition, so the same missed match extinguishes both stacked
    penalties and zeroes the counter.
\end{itemize}

\section{Summary}

Every relation in the refined schema is in 3NF, and 12 of 13 are in
BCNF. Club is kept in 3NF by choice: the sole BCNF-violating dependency
(\texttt{stadium\_ID} $\to$ \texttt{city}) is controlled by triggers so
that the redundancy it introduces is always consistent with Stadium,
and decomposing would shift the enforcement burden without a real
integrity benefit. All business rules not expressible by
\texttt{CHECK} are implemented as triggers or stored procedures, giving
the database independent authority over data integrity as required by
the project specification.

\end{document}
